Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-04-01 Grassmannian_cells.tex
Lemma 2.1, Lemma 2.7, Proposition 2.1 changed
Added argument concerning irreducible morphism to Lemma 2.1
Added Lemma 2.7 which should describe the empty fibers (details have to
be checked)
Changed Proposition 2.1 (2) under consideration of 2.7

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym}
\usepackage[margin=1in]{geometry}

\input xy
\xyoption{all}

@@

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}

@@

\begin{lemma}
  For any $m\ge1$, we have 
  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0.\]  
  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
\end{lemma}
Write $P_{m+1}^{\{f_1,\ldots,f_k\}}$ for the cokernel of the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$, i.e. we have a short exact sequence
\begin{equation}
\label{eq:truncated preprojectives}
  0\longrightarrow P_m^k\stackrel{[f_1\ \cdots\ f_k]}{\longrightarrow} P_{m+1}\longrightarrow P_{m+1}^{\{f_1,\ldots,f_k\}}\longrightarrow 0.
\end{equation}
\begin{lemma}
  For $m\ge1$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the following hold:
  \begin{enumerate}
    \item $\Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})$ is naturally isomorphic to $\Hom(P_m,P_{m+1})/\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ is naturally isomorphic to $\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Hom(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)=0$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Applying the functor $\Hom(P_m,-)$ to the sequence \eqref{eq:truncated preprojectives} gives the exact sequence
  \[0\longrightarrow \Hom(P_m,P_m^k)\stackrel{[f_1\ \cdots\ f_k]\circ}{\longrightarrow} \Hom(P_m,P_{m+1})\longrightarrow \Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})\longrightarrow 0,\]
  establishing (1).

@@

\begin{lemma}
  \begin{enumerate}
    \item $\tilde P_m^{(n-1)}\cong\tilde P_{m-1}^{(1)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}

\begin{corollary}
  observe when fibers are empty
\end{corollary}

\begin{proposition}
  Consider $\psi:Gr^{\tilde Q}_\tbfe(\tilde P_m)\to\bigsqcup_{\tbff+\tbfg=\tbfe} Gr^{\tilde Q}_\tbff(\bigoplus_{i=1}^k \tilde P_{m-1,i})\times Gr^{\tilde Q}_\tbfg(\tilde P_m^{(k)}$.  Then the following hold:
  \begin{enumerate}
    \item For $V\subsetneq \tilde P_m^{(k)}$ and $U\subset\bigoplus_{i=1}^k\tilde P_{m-1,i}$, we have $\psi^{-1}(U,V)=\AA^{\langle V,\bigoplus_{i=1}^k\tilde P_{m-1,i}/U\rangle}$.
    \item If $V=\tilde P_m^{(k)}$ and the fiber is not empty, then $\psi^{-1}(U,V)$ is constant.
  \end{enumerate}
\end{proposition}
\begin{proof}
  (1) $V$ is preprojective but $\bigoplus_{i=1}^k\tilde P_{m-1,i}/U$ is not unless $U=0$

  (2) \[0\longrightarrow[V,P/U]\longrightarrow[V,U]^1\longrightarrow[V,P]^1\longrightarrow[V,P/U]^1\longrightarrow0\]
      and the middle map is surjective.
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
\begin{thebibliography}{10}
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\end{thebibliography}

>>>>

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym}
\usepackage[margin=1in]{geometry}
\usepackage{extarrows}
\input xy
\xyoption{all}

@@

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}

@@

\begin{lemma}
  For any $m\ge1$, we have 
  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0.\]  
  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
\end{lemma}
\begin{proof} Note that we have a Auslander-Reiten sequence
\[0\longrightarrow P_{m-1}\longrightarrow P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0\]
with irreducible homomorphisms $f_i$ which are also linear independent. Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. If we pick $k<n$ linear independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. Indeed otherwise, at least one of the homomorphisms were forced to factor through a representation $Z\neq P_{m}$.

 Write $P_{m+1}^{\{f_1,\ldots,f_k\}}$ for the cokernel of the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$, i.e. we have a short exact sequence
\begin{equation}
\label{eq:truncated preprojectives}
  0\longrightarrow P_m^k\stackrel{[f_1\ \cdots\ f_k]}{\longrightarrow} P_{m+1}\longrightarrow P_{m+1}^{\{f_1,\ldots,f_k\}}\longrightarrow 0.
\end{equation}
\end{proof}
\begin{lemma}
  For $m\ge1$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the following hold:
  \begin{enumerate}
    \item $\Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})$ is naturally isomorphic to $\Hom(P_m,P_{m+1})/\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ is naturally isomorphic to $\kk f_1\oplus\cdots\oplus\kk f_k$;
    \item $\Hom(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)=0$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  Applying the functor $\Hom(P_m,-)$ to the sequence \eqref{eq:truncated preprojectives} gives the exact sequence
  \[0\longrightarrow \Hom(P_m,P_m^k)\stackrel{[f_1\ \cdots\ f_k]\circ}{\longrightarrow} \Hom(P_m,P_{m+1})\longrightarrow \Hom(P_m,P_{m+1}^{\{f_1,\ldots,f_k\}})\longrightarrow 0,\]
  establishing (1).

@@

\begin{lemma}
  \begin{enumerate}
    \item $\tilde P_m^{(n-1)}\cong\tilde P_{m-1}^{(1)}$
    \item The subrepresentation $\bigoplus_{i=1}^{k-1}\tilde P_{m-1,i}\oplus\bigoplus_{i=1}^k\tilde P_{m-2,i}\subset\bigoplus{i=1}^k\tilde P_{m-1,i}$ is in $(\tilde P_m^{(k)})^\perp$ and
      \[\Ext(\bigoplus_{i=1}^k\tilde P_{m-1,i},\tilde P_m^{(k)})\cong\Ext(\tilde P_{m-1}^{(k)},\tilde P_m^{(k)})\]
      where $\tilde P_{m-1}^{(k)}$ above denotes the cokernel of the inclusion.
  \end{enumerate}
\end{lemma}

\begin{lemma}\label{quotient}
 Consider $\ses{\tilde P_m}{\tilde P_{m+1}^{(k)}}{\tilde P_{m+1}^{(k-1)}}$. If $U\subset P_m$ such that $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$, then we have $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$.
\end{lemma}
\begin{proof}
First observe that $\tilde P_m\in ~^\perp \tilde P_{m+1}^{(k-1)}$ and $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$. As $U$ is preprojective with $U\subset \tilde P_m\subset \tilde P_{m+1}^{(k)}$, it follows that $\Hom(U,\tilde P_{m+1}^{(k-1)})=0$ (check again). This yields $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})=0$. Consider
\[0\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]\longrightarrow[\tilde P_{m+1}^{(k-1)},U]^1\xlongrightarrow{\pi}[\tilde P_{m+1}^{(k-1)},\tilde P_m]^1\longrightarrow[\tilde P_{m+1}^{(k-1)},\tilde P_m/U]^1\longrightarrow 0.\]
Since $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)=k$, we have that $\pi$ is surjective if $\pi\neq 0$. If $U=U_1\oplus\ldots\oplus U_r$ is a decomposition into indecomposable representations, we can write $\pi=(\pi_1,\ldots,\pi_r)$ and it suffices to show that $\pi_j\neq 0$ for at least one $j=1,\ldots,r$. Thus we can without loss of generality assume that $U$ is indecomposable (check again). This means that $U\cong \tilde P_{i}$ for some $i=1,\ldots,m$. Moreover, $\tilde P_{m+1}/U$ is indecomposable (details).

If $\tilde P_{m+1}/U$ is exceptional, $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})=0$ together with \cite[Theorem 4.1]{sch} yields that either $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)=0$ or $\Hom(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)=0$. Thus $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)\neq 0$ would imply that $\Ext(\tilde P_{m+1}^{(k-1)},U)=0$. But this contradicts the assumption and the claim follows.

Thus assume that $\tilde P_{m+1}/U$ has self-extensions. Since $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})= 0$, from \cite[Lemma 4.1]{hr} we obtain that every homomorphism $\tilde P_{m+1}^{(k-1)}\to \tilde P_m/U$ is injective or surjective. By a dimension count it needs to be surjective (case $k=1$?). Such a homomorphism would yield $\Ext(\tilde P_m/U,\tilde P_{m+1}^{(k-1)})\twoheadrightarrow\Ext(\tilde P_m/U,\tilde P_m/U)\neq 0$ which contradicts the further observations. Thus $\Ext(\tilde P_{m+1}^{(k-1)},U)\neq 0$, already yields $\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m/U)= 0$ in this case.

\end{proof}

\begin{proposition}
  Consider $\psi:Gr^{\tilde Q}_\tbfe(\tilde P_m)\to\bigsqcup_{\tbff+\tbfg=\tbfe} Gr^{\tilde Q}_\tbff(\bigoplus_{i=1}^k \tilde P_{m-1,i})\times Gr^{\tilde Q}_\tbfg(\tilde P_m^{(k)}$.  Then the following hold:
  \begin{enumerate}
    \item For $V\subsetneq \tilde P_m^{(k)}$ and $U\subset\bigoplus_{i=1}^k\tilde P_{m-1,i}$, we have $\psi^{-1}(U,V)=\AA^{\langle V,\bigoplus_{i=1}^k\tilde P_{m-1,i}/U\rangle}$.
    \item If $V=\tilde P_m^{(k)}$, the fibre $\psi^{-1}(U,V)$ is non-empty and of constant dimension if and only if $\Ext(\tilde P_m^{(k)},U)\neq 0$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  (1) $V$ is preprojective but $\bigoplus_{i=1}^k\tilde P_{m-1,i}/U$ is not unless $U=0$

  (2) 
If $\Ext(\tilde P_m^{(k)},U)= 0$, the fibre is empty because the sequence is non-split. If $\Ext(\tilde P_m^{(k)},U)\neq 0$, Lemma \ref{quotient} yields $\Ext(\tilde P_m^{(k)},\tilde P_{m-1}/U)=0$. In particular, the dimension is constant if it is not empty and
$\Ext(\tilde P_{m+1}^{(k-1)},U)\xlongrightarrow{\pi}\Ext(\tilde P_{m+1}^{(k-1)},\tilde P_m)$ is surjective. But this already means that the fibre is not empty.
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Compatible Pairs Label Cells in $Gr^Q_\bfe(P_m)$}
\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{hr} Happel, D., Ringel, C.M.: Tilted Algebras. Transactions of the American Mathematical Society {\bf 274}, no.2, 399-443 (1982).

\bibitem{sch} Schofield, A.: General representations of quivers. Proceedings of the London Mathematical Society (3) \textbf{65}(1), 46-64 (1992).

\end{thebibliography}
